---
title: "Burkina Faso Weather Forecast"
author: "Derek Morr"
date: "5/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries, message=FALSE, warning=FALSE, include=FALSE}
library(terra)
library(tmap)
library(stars)
library(glue)
library(purrr)
library(dplyr)
```

# Constants

First, we define some constants.

```{r}
output_dir <- "c:/users/derek/Documents/burkina-forecat-map/output"
adm1_file <- "shp/bfa_admbnda_adm1_igb_20200323.shp"
adm2_file <- "shp/bfa_admbnda_adm2_igb_20200323.shp"
adm3_file <- "shp/bfa_admbnda_adm3_igb_20200323.shp"
prov_capitals_file <- "new_data/Bf_prov.shp"
forecast_input_tiff <- "plot_ensemble_median_total_rainfall.tiff"
output_filename <- "bf_rainfall_forecast_2022-05-26-to-2022-06-26.png"
```

Make the output directory, if necessary.

```{r}
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}
```


# Crop the GeoTIFF

First, we crop the forecast GeoTIFF to Burkina Faso. We use the 
[R terra package](https://rspatial.github.io/terra/) to do this.

```{r}
forecast <- rast(forecast_input_tiff, lyrs = 1)
adm1 <- vect(adm1_file)
adm2 <- vect(adm2_file)
adm3 <- vect(adm3_file)
prov_capitals <- vect(prov_capitals_file)

# crop the forecast GeoTIFF to BFA boundaries
bfa_forecast <- terra::mask(forecast, adm1, touches = TRUE)
bfa_forecast <- terra::crop(bfa_forecast, adm1, snap = "out")
```

# TMAP

The [tmap package](https://r-tmap.github.io/tmap/) has many tools for making 
beautiful maps programatically.

## Whole Country Map

A note on styling. I chose a built-in blue-scale color palette and quantile
styling. This was a quick way to approximate the original map from Romaric.
If this is not acceptable, we can adjust the `style = "quantile"` argument 
to adjust the raster styling. We can also manually set breaks and a palette.

```{r}
country_map <- tm_shape(st_as_stars(bfa_forecast)) + 
  tm_raster(
    title = "Quantités de pluie (mm)",
    palette = "Blues",
    style = "quantile"
  ) +
  tm_shape(st_as_sf(prov_capitals)) +
  tm_borders(col = "black") +
  tm_text(
    text = "CHF_LIEU",
    col = "red",
    size = 0.6
  ) +
  tm_compass(
    type = "rose",
    show.labels = 2,
    position = c("left", "top"),
    size = 4
  ) +
  tm_graticules(alpha = 0) +
  tm_layout(
    main.title = "Total pluviométrique\ndu 26 Mai 2022 au 26 Juin 2022",
    main.title.fontface = "bold",
    main.title.position = "center",
    legend.position = c("right", "bottom"),
    inner.margins = 0.06
  )
```

Manually inspect the map:

```{r}
country_map
```

Save it to a file:

```{r}
tmap_save(country_map, glue("{output_dir}/{output_filename}"))
```


# Per-region maps

```{r}
mk_region_map <- function(region) {
  region_vect <- adm3[adm3$ADM1_FR == region]
  
  # Sahel needs special handling for cropping & masking.
  if (region != "Sahel") {
    input_raster <- forecast
  } else{
    input_raster <- bfa_forecast
  }

  region_raster <- terra::mask(input_raster, region_vect, touches = TRUE)
  region_raster <- terra::crop(region_raster, region_vect, snap = "out")
  
  special_compass_regions <- c(
    "Centre", "Centre-Est", "Centre-Ouest", "Hauts-Bassins"
  )
  if (region %in% special_compass_regions) {
    compass_position <- c("right", "top")
  } else {
    compass_position <- c("left", "bottom")
  }
  #compass_position <- if_else(
  #  region %in% special_compass_regions, 
  #  c("right", "top"),
  #  c("left", "bottom")
  #)
  
  special_scale_bar_regions <- c(
    "Cascades", "Centre", "Est", "Hauts-Bassins", "Nord"
  )
  if (region %in% special_scale_bar_regions) {
    scale_bar_position <- c("right", "bottom")
  } else {
    scale_bar_position <- c("center", "bottom")
  }
  #scale_bar_position <- if_else(
  #  region %in% special_scale_bar_regions,
  #  c("right", "bottom"),
  #  c("center", "bottom")
  #)
  
  map <- tm_shape(st_as_stars(region_raster)) + 
    tm_raster(
      title = "Quantités de pluie (mm)",
      palette = "Blues",
      style = "quantile"
    ) +
    tm_shape(st_as_sf(region_vect)) +
    tm_borders(col = "black") +
    tm_text(
      text = "ADM3_FR",
      col = "red",
      size = 0.5
    ) +
    tm_compass(
      type = "rose",
      show.labels = 2,
      position = compass_position,
      size = 2
    ) +
    tm_scale_bar(position = scale_bar_position) +
    tm_graticules(alpha = 0) +
    tm_layout(
      main.title = glue("Total pluviométrique en {region}\n26 Mai 2022 au 26 Juin 2022"), 
      main.title.fontface = "bold",
      main.title.position = "center",
      legend.outside.position = c("right", "center"),
      legend.outside = TRUE,
      inner.margins = 0.1
    )
  
  map
}
```

```{r}
regions <- sort(unique(adm3$ADM1_FR))
region_maps <- regions |> map(mk_region_map)
```

broken regions:

Sahel, 

```{r}
for (i in 1:length(regions)) {
  region <- regions[i]
  map <- region_maps[[i]]
  print(glue("writing {region} map"))
  tmap_save(map, glue("{output_dir}/{region}.png"))
}
```

